Seagrass mapping and monitoring \\ along the coasts of Crete, Greece

### 2011-02-17 MScDiplom_Lemenkova.tex
Comment 32 of Valentijn (final): updated Research Question 2, proved hypothesis

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{Sensors.jpg}
	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths, as stripes shown spectral bands covered by these sensors. Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-33}$>$}
	\label{fig:4.6}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.50]{GNU-13.jpg}
		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}. Series 1-100. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot}­. Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-41}$>$}
		\label{fig:4.13}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.5]{GNU-14.jpg}
		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}. Series 126-150. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization. Page: $<$\pageref{page-39}$>$­}
		\label{fig:A.27}
	\end{center}
\end{figure}

@@

\subsubsection{Remote sensing application}
The \textit{in situ} spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
types as the different sensors would percept them: \ac{MODIS}\index[concepts]{Remote sensing!multispectral imagery!MODIS}, \ac{ASTER}\index[concepts]{Remote sensing!multispectral imagery!ASTER}, \ac{MERIS}\index[concepts]{Remote sensing!multispectral imagery!MERIS}, \ac{SeaWiFS}\index[concepts]{Remote sensing!multispectral imagery!SeaWiFS} and \ac{CZCS}\index[concepts]{Remote sensing!multispectral imagery!CZCS}.\\
These sensors vary in technical characteristics and therefore, have different spectral sensitivity, 
which we briefly illustrated in a small summary table (Tab.\ref{tab:1})\label{page-intmap}
These models simulate spectral views of the chosen sensors, how these sensors will ``see" seagrass (Fig.\ref{fig:28}) and sand as
pixels, with accepted default atmosphere and water column effects (given by \ac{WASI} software). In such
a way we defined an empirical upper limit to the discriminative potential of these sensors.

@@

In \textit{Chapter 3}: \nameref{sec:3}, describing data collection, a videographic approach tested in previous works, has been applied during the summer fieldwork, when we captured imagery and video footage of the seafloor on several routes of the boat in the Ligaria beach.\\

@@

10 \% reflectance in spectra of 500-600 nm, while sand has spectral reflectance approaching 33\% in its highest values. These results indicate that seagrass \textit{P.oceanica} can in general be detected and discriminated from other seafloor cover types with varying environmental conditions, i.e. water column height, by hyperspectral spectroradiometers (\ac{Trios-RAMSES}), which positively answers the \textbf{first research question} of this thesis (``Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying in-situ conditions ?"). \vspace{1ex}\\
Further in \textit{Chapter 4}, \nameref{sec:4}, studies of the broadband versus narrowband sensors demonstrate that simulated spectra of the seagrass, made using \ac{WASI} modeller, have the best results at \ac{CZCS} scanner, especially devoted to the measurement of ocean color. The spectrum of \textit{P.oceanica} reflectance, simulated for \ac{CZCS}, covers the wavelength interval of 400-800 nm, and is distinctive for various depths. Other remote sensors (\ac{MODIS}, \ac{SeaWiFS}) may also be used for the seagrass mapping, because their technical characteristics enable to spectrally discriminate \textit{P.oceanica} seagrass from other seafloor cover types (Fig.\ref{fig:4.6}), particularly carbonate sand as tested in the current work. Therefore, the \textbf{second research question} of this MSc thesis (``Can broadband and hyperspectral remote sensing data be used for the mapping of \textit{P.oceanica}: what sensor yields enough radiometric details to discriminate seagrass from other seafloor cover types?") is answered with ``yes'', and the most suitable sensor is the Coastal Zone Color Scanner \ac{CZCS}.\vspace{1ex}\\

@@

\begin{table}[H]\footnotesize
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}. \ac{SPSS}. Page: $<$\pageref{page-reg}$>$}
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline\hline
			\textbf{R} & \textbf{R square} & \textbf{Adjusted R square} & \textbf{Std. Error of the Estimate} \\ \hline\hline
			.277 1 & .077 & .068126692 & .462 \\ \hline
		\end{tabular}
	\end{center}
	The independent variable is wavelength. 
	\label{tab:5}
\end{table}

\begin{table}[H]\footnotesize
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}. \ac{SPSS}. Page: $<$\pageref{page-reg}$>$}
	\begin{center}
		\begin{tabular}{|c|c|c|c| c|c|}
			\hline\hline
			& \textbf{Sum of squares} & \textbf{df} & \textbf{Mean Square} & F & Sig.\\ \hline\hline
			Regression & 1.810 & 1 & 1.810 & 8.477 & .004\\ \hline
			Residual & 21.781 & 102 & .214 & & \\ \hline
			Total & 23.592 & 103 & & & \\ \hline
		\end{tabular}
	\end{center}
	The independent variable is wavelength. 
	\label{tab:6}
\end{table}

\begin{table}[H]\footnotesize
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textit{P.oceanica}. \href{http://www.spss.com/}{SPSS}. Page: $<$\pageref{page-reg}$>$}
	\begin{center}
		\begin{tabular}{| c | c | p{2cm}| c | c | c |}
			\hline\hline
			& \multicolumn{2}{|c|}{Unstandardized Coefficients}\\
			\cline{2-3}
			& B & Std.Error & Stand. Coef. Beta & t & Sig.\\ \hline\hline
			wavelength & -.001 & .000 & -.277 & -2.912 & .004 \\ \hline
			constant & .464 & .121 & & 3.821 & .000 \\ \hline
		\end{tabular}
	\end{center}
	The underlying process assumed is independence (white noise).\\
	 Based on the asymptotic chi-square approximation.
	\label{tab:7}
\end{table}

@@

\begin{figure}[H]
	\centering
	\subfloat [CZCS]{\label{fig:29a}\includegraphics[width=0.4\textwidth]{Fig-35a.jpg}}
	\subfloat [MERIS]{\label{fig:29b}\includegraphics[width=0.4\textwidth]{Fig-35b.jpg}}\\
	\subfloat [MODIS]{\label{fig:29c}\includegraphics[width=0.4\textwidth]{Fig-35c.jpg}}
	\subfloat [SeaWiFS]{\label{fig:29d}\includegraphics[width=0.4\textwidth]{Fig-35d.jpg}}
	\caption{WASI-simulated remote sensing reflectance of \textit{P.oceanica} at various sensors. Page: $<$\pageref{page-33}$>$}
	\label{fig:4.6W}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.35]{GNU-11.jpg}
		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display. Pages: $<$\pageref{page-21}$>$, $<$\pageref{dataprep}$>$­}
		\label{fig:A.25}
	\end{center}
\end{figure}
\pagebreak

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.35]{GNU-12.jpg}
		\caption{Remote sensing reflectance of \textit{P.oceanica}. Series 1-25. \\Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range. \\ \href{http://www.gnuplot.info/}{Gnuplot}­. Page: $<$\pageref{page-21}$>$}
		\label{fig:A.26}
	\end{center}
\end{figure}

>>>>

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{Sensors.jpg}
	\caption{Combined plots of simulated remote sensing reflectance of \textit{P.oceanica} at various sensors: \ac{MODIS}, \ac{MERIS}, \ac{SeaWiFS} and \ac{CZCS}, iterated over three depths, as stripes shown spectral bands covered by these sensors. Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-22}$>$, $<$\pageref{page-33}$>$, $<$\pageref{secondq}$>$}
	\label{fig:4.6}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.50]{GNU-13.jpg}
		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}. Series 1-100. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot}­. Pages: $<$\pageref{page-21}$>$, $<$\pageref{page-41}$>$, $<$\pageref{secondq}$>$}
		\label{fig:4.13}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.5]{GNU-14.jpg}
		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}. Series 126-150. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization. Pages: $<$\pageref{page-39}$>$­, $<$\pageref{secondq}$>$­}
		\label{fig:A.27}
	\end{center}
\end{figure}

@@

\subsubsection{Remote sensing application}
The \textit{in situ} spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
types as the different sensors would percept them: \ac{MODIS}\index[concepts]{Remote sensing!multispectral imagery!MODIS}, \ac{ASTER}\index[concepts]{Remote sensing!multispectral imagery!ASTER}, \ac{MERIS}\index[concepts]{Remote sensing!multispectral imagery!MERIS}, \ac{SeaWiFS}\index[concepts]{Remote sensing!multispectral imagery!SeaWiFS} and \ac{CZCS}\index[concepts]{Remote sensing!multispectral imagery!CZCS}.\\
These sensors vary in technical characteristics and therefore, have different spectral sensitivity, 
which we briefly illustrated in a small summary table (Tab.\ref{tab:1})\label{page-intmap}
These models simulate spectral views of the chosen sensors, how these sensors will ``see" seagrass (Fig.\ref{fig:28}) and sand as
pixels, with accepted default atmosphere and water column effects (given by \ac{WASI} software). In such
a way we defined an empirical upper limit to the discriminative potential of these sensors. 
The analysis of the remote sensing reflectance simulated by these sensors shows that the measured spectra of seagrass \textit{P.oceanica} (Fig.\ref{fig:4.13}, Fig.\ref{fig:A.27}) were statistically different at most of the spectral bands (Fig.\ref{fig:A.25}, Fig.\ref{fig:A.26}), with a 95\% confidence level (p value $< 0. 05$).  The F values of the test (F=8.477 as on Tab.\ref{tab:6}) are greater than F critical value, which is 2.64 at 0.05 confidence level. This proves that broadband and hyperspectral sensors enable spectral discrimination of seagrass, i.e. $H_o: \mu1\neq\mu2\neq\mu3\neq ... \neq\mu n$. Therefore, we answer the second research question positively, i.e. broadband and hyperspectral sensors (Fig.\ref{fig:4.6} and Fig.\ref{fig:4.6W}) provide enough radiometric information for spectral discrimination of seagrass and can therefore be used for mapping of seagrass \textit{P.oceanica}.\label{secondq}

@@

In \textit{Chapter 3}: \nameref{sec:3}, describing data collection, a videographic approach tested in previous works, has been applied during the summer fieldwork, when we captured imagery and video footage of the seafloor on several routes of the boat in the Ligaria beach.\vspace{1ex}\\

@@

10 \% reflectance in spectra of 500-600 nm, while sand has spectral reflectance approaching 33\% in its highest values. These results indicate that seagrass \textit{P.oceanica} can in general be detected and discriminated from other seafloor cover types (carbonate sand) with varying environmental conditions, i.e. water column height, by hyperspectral spectroradiometers (\ac{Trios-RAMSES}), which positively answers the \textbf{first research question} of this thesis (``Is \textit{P.oceanica} spectrally distinct from carbonate sand with varying \textit{in-situ} conditions ?").
Further in \textit{Chapter 4}, \nameref{sec:4}, studies of the broadband and narrowband sensors demonstrate that simulated spectra of the seagrass, made using \ac{WASI} modeller, have the best results at \ac{CZCS} scanner, especially devoted to the measurement of ocean color. The spectrum of \textit{P.oceanica} reflectance, simulated for \ac{CZCS}, covers the wavelength interval of 400-800 nm, and is distinctive for various depths. Other remote sensors (\ac{MODIS}, \ac{SeaWiFS}) may also be used for the seagrass mapping, because their technical characteristics enable to spectrally discriminate \textit{P.oceanica} seagrass from other seafloor cover types (Fig.\ref{fig:4.6}), particularly carbonate sand as tested in the current work. Therefore, the \textbf{second research question} of this MSc thesis (``Do broadband and hyperspectral sensors provide enough radiometric information for spectral discrimination of seagrass, 
and therefore, can be used for mapping of \textit{P.oceanica} ?") is answered with ``yes'', and the most suitable sensor is the Coastal Zone Color Scanner \ac{CZCS}.\vspace{1ex}\\

@@

\begin{table}[H]\footnotesize
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}. \ac{SPSS}. Page: $<$\pageref{page-reg}$>$}\label{tab:5}
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline\hline
			\textbf{R} & \textbf{R square} & \textbf{Adjusted R square} & \textbf{Std. Error of the Estimate} \\ \hline\hline
			.277 1 & .077 & .068126692 & .462 \\ \hline
		\end{tabular}
	\end{center}
	The independent variable is wavelength. 	
\end{table}

\begin{table}[H]\footnotesize
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}. \ac{SPSS}. Pages: $<$\pageref{page-reg}$>$, $<$\pageref{secondq}$>$}\label{tab:6}
	\begin{center}
		\begin{tabular}{|c|c|c|c| c|c|}
			\hline\hline
			& \textbf{Sum of squares} & \textbf{df} & \textbf{Mean Square} & F & Sig.\\ \hline\hline
			Regression & 1.810 & 1 & 1.810 & 8.477 & .004\\ \hline
			Residual & 21.781 & 102 & .214 & & \\ \hline
			Total & 23.592 & 103 & & & \\ \hline
		\end{tabular}
	\end{center}
	The independent variable is wavelength. 	
\end{table}

\begin{table}[H]\footnotesize
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textit{P.oceanica}. \href{http://www.spss.com/}{SPSS}. Page: $<$\pageref{page-reg}$>$}\label{tab:7}
	\begin{center}
		\begin{tabular}{| c | c | p{2cm}| c | c | c |}
			\hline\hline
			& \multicolumn{2}{|c|}{Unstandardized Coefficients}\\
			\cline{2-3}
			& B & Std.Error & Stand. Coef. Beta & t & Sig.\\ \hline\hline
			wavelength & -.001 & .000 & -.277 & -2.912 & .004 \\ \hline
			constant & .464 & .121 & & 3.821 & .000 \\ \hline
		\end{tabular}
	\end{center}
	The underlying process assumed is independence (white noise).\\
	 Based on the asymptotic chi-square approximation.	
\end{table}

@@

\begin{figure}[H]
	\centering
	\subfloat [CZCS]{\label{fig:29a}\includegraphics[width=0.4\textwidth]{Fig-35a.jpg}}
	\subfloat [MERIS]{\label{fig:29b}\includegraphics[width=0.4\textwidth]{Fig-35b.jpg}}\\
	\subfloat [MODIS]{\label{fig:29c}\includegraphics[width=0.4\textwidth]{Fig-35c.jpg}}
	\subfloat [SeaWiFS]{\label{fig:29d}\includegraphics[width=0.4\textwidth]{Fig-35d.jpg}}
	\caption{WASI-simulated remote sensing reflectance of \textit{P.oceanica} at various sensors. Pages: $<$\pageref{page-33}$>$, $<$\pageref{secondq}$>$}
	\label{fig:4.6W}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.35]{GNU-11.jpg}
		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display. Pages: $<$\pageref{page-21}$>$, $<$\pageref{dataprep}$>$­, $<$\pageref{secondq}$>$}
		\label{fig:A.25}
	\end{center}
\end{figure}
\pagebreak

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.35]{GNU-12.jpg}
		\caption{Remote sensing reflectance of \textit{P.oceanica}. Series 1-25. \\Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range. \\ \href{http://www.gnuplot.info/}{Gnuplot}­. Page: $<$\pageref{page-21}$>$, $<$\pageref{secondq}$>$}
		\label{fig:A.26}
	\end{center}
\end{figure}
